---
title: "Carbonate Chemistry"
author: "Jaydee Sereewit"
date: "11/5/2021"
output: pdf_document
---

# Installing older seacarb package and OTools package
Uncomment lines 13 and 14 if you need to install those packages.  
```{r Libraries, include=FALSE}
library(remotes)
library(devtools)
#install_version("seacarb", version = "3.2.14", repos = "http://cran.us.r-project.org")
#install_github("KirtOnthank/OTools")
library(seacarb)
library(OTools)
library(readODS)
```

# Calculating pH from spectrophotometric data
## Getting file names
```{r Reading in filenames}
files=list.files(recursive=T)
```

## Selecting only the pH data
```{r Selecting pH data}
files.ph=files[grep("Spec_pH_Data",files)]
```

## Running the first dataset
First we extract the date form the filename.
```{r Extracting pH date}
date.ph=gsub(".*(\\d\\d)_(\\d\\d)_(\\d\\d).ods","\\1-\\2-\\3",files.ph[1])
```

Then read in the dataset.
```{r Reading in pH dataset}
data.ph=read_ods(files.ph[1],col_names=T)
```

Now, we calculated the pH of all measurements in the file with the specpH() function.
```{r Calculating pH}
ph=specpH(data.ph$S,data.ph$T,data.ph$A434,data.ph$A578,data.ph$A730,data.ph$B434,
          data.ph$B578,data.ph$B730)
```

Next, we gets the mean pH calculated for each tank.
```{r Calculating pH means}
ph.means=aggregate(ph~data.ph$Tank,FUN="mean")
```

Finally, do some formatting.
```{r pH formatting}
colnames(ph.means)[1]="tank"
ph.means$date=date.ph
ph.means$S=aggregate(data.ph$S~data.ph$Tank,FUN="mean")[,2]
carbchem=ph.means
```

Repeating the process with each of the following datasets and adding them to he dataset.
```{r pH calculation loop}
for (i in 2:length(files.ph)){
  date.ph=gsub(".*(\\d\\d)_(\\d\\d)_(\\d\\d).ods","\\1-\\2-\\3",files.ph[i])
  data.ph=read_ods(files.ph[i],col_names=T)
  ph=specpH(data.ph$S,data.ph$T,data.ph$A434,data.ph$A578,data.ph$A730,
            data.ph$B434,data.ph$B578,data.ph$B730)
  ph.means=aggregate(ph~data.ph[,1],FUN="mean")
  colnames(ph.means)[1]="tank"
  ph.means$date=date.ph
  ph.means$S=aggregate(data.ph[,3]~data.ph[,1],FUN="mean")[,2]

  carbchem=rbind(carbchem,ph.means)
}
```

# Calculating alkalinity from titration data

Similar to how we did with the pH data, we extract the tank # from the filename, then read in the data and calculate the alkalinity from the data.
Finally, we will organize it into a dataframe called "alk".
```{r First titration calculation}
files.ti=files[grep("Titration_data/Titration",files)]

ti.tank=gsub(".*Tank (\\d+) Run \\d.ods","\\1",files.ti[1])
ti.filename=gsub(".*Titration_data/(.*)\\.ods","\\1",files.ti[1])
ti=read_ods(files.ti[1],col_names = T)
ti.at=at(ti[1,1],ti[1,8],ti[1,2],1,ti[1,3],ti[1,4],ti[1,5],ti[,6],ti[,7])[1]
alk=data.frame(filename=ti.filename,tank=ti.tank,alk=ti.at)
```

Then, we repeat for the rest of the data in the folder.
```{r Titration calculation loop}
for (i in 2:length(files.ti)){
  ti.tank=gsub(".*Tank (\\d+) Run \\d.ods","\\1",files.ti[i])
  ti.filename=gsub(".*Titration_data/(.*)\\.ods","\\1",files.ti[i])
  ti=read_ods(files.ti[i],col_names = T)
  ti.at=at(ti[1,1],ti[1,8],ti[1,2],1,ti[1,3],ti[1,4],ti[1,5],ti[,6],ti[,7])[1]
  alk=rbind(alk,data.frame(filename=ti.filename,tank=ti.tank,alk=ti.at))
}
```

Now, we get the alkalinity mean for each tank.
```{r Alkalinity means}
alk.means=aggregate(alk~tank,data=alk,FUN="mean")
```

We take the mean alkalniity for each tank and put it into the carbchem dataset with the pH values.
```{r Inserting alkalinity into carbchem}
carbchem=carbchem[carbchem$tank %in% alk.means$tank,]
carbchem$alk=0

for (i in 1:nrow(carbchem)){
  carbchem$alk[i]=alk.means$alk[alk.means$tank==carbchem$tank[i]]
}

```

# Calculating pCO2 from the data
```{r pCO2 calculation}
carbchem$pco2=carb(8,carbchem$ph,carbchem$alk,carbchem$S,10.8)$pCO2

```


# Writing the data out to a .cvs file
```{r Writing data out}
write.csv(carbchem,"carbonate_chemistry.csv",row.names = F)
```

# The final dataset
```{r}
carbchem
```







